An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-06-07 EconGrowth.Rnw
Line shortening.

Values of $\varepsilon_i$ and $\phi_i$ were calculated from the REXS database supplied by \citet{Warr:2012a}. Values of $X$ were calculated by methods discussed in Section~\ref{sec:X_Data}. For most countries, values of $\phi_i$ are not constant with time, as economic sectors with different distributions of raw exergy inputs grow at different rates. \citeapos{Warr:2012a} REXS database contains information for developed economies (United States, United Kingdom, and Japan) from 1900--2000, and we used data from the 1980--2000 timeframe. Extending the REXS database to the BRICS, petrol, and LDC economies is beyond the scope of this work, because longitudinal data for the efficiency of different end uses ($\varepsilon_i$) and the distribution of exergy sources to end uses ($\phi_i$) are extremely difficult to find.

@@

Each of the economic growth models discussed in Section~\ref{sec:Prod_Func_Forms} has parameters which must be estimated for each economy (see Section~\ref{sec:Economy_Types}) and each energy quantification (see Section~\ref{sec:Quantification_of_Energy}) via least-squares methods. The following subsections discuss parameter estimation methods interpret the results.

@@

% \citet{Henningsen:2011td}, in their detailed analysis of \citet{Kemfert:1998vd}, found that 
% 
% \begin{quote}
% \ldots the Levenberg-Marquardt and the PORT algorithms are---at least in this study---most likely to find the coefficients that give the best fit to the model, where the PORT algorithm can be used to restrict the estimates to the economically meaningful region.
% \end{quote}

@@

\begin{itemize}
  \item{The use of chemical exergy, as opposed to thermal energy, in economic growth models is probably unnecessary. Table~\ref{tab:Exergy_Allocation} shows that exergy allocation factors are all close to unity. Therefore, there is a small difference in absolute magnitude between $Q$ and $X$. However, there is no difference in indexed magnitude between $q$ and $x$, unless there are large changes in energy mix for an economy. Figure~\ref{fig:Historical_Lattice_Graph} shows that $q$ and $x$ lines lie atop one another for almost every economy.}
  \item{The concept of useful work ($u$) is potentially useful and should be explored beyond developed economies. Figure~\ref{fig:Historical_Lattice_Graph} shows that $u$ grows at very different rates from thermal energy ($q$) or exergy ($x$). Useful work may contain different information from thermal energy and exergy.}
  \item{Figure~\ref{fig:CDeGridLattice} shows that the set of parameters ($\lambda$, $\alpha$, $\beta$, and $\gamma$) that minimizes $SSE$ for the energy-augmented Cobb-Douglas production function always (with the exception of Saudi Arabia) has $\beta = 0$ or $\gamma = 0$. This result implies that it is unnecessary to include both labor ($l$) and energy ($e$) in the Cobb-Douglas production function.}
  \item{Partial residual plots for both the Cobb-Douglas model (Figures~\ref{fig:CDPartialResidualsQ}--\ref{fig:CDPartialResidualsU} and CES model (Figures~\ref{fig:CESPartialResidualsQ}--\ref{fig:CESPartialResidualsU}) indicate that augmenting the Cobb-Douglas and CES production functions with energy is not beneficial.}
  \item{Figure~\ref{fig:CIvsParam_Lattice_Graph} shows that important energy parameters for both the energy-augmented Cobb-Douglas and CES production functions ($\gamma$ and $\delta$, respectively) have large confidence intervals such that we are unable to say (at the 95\% confidence level) that energy is an important factor for economic growth, vis-\`{a}-vis capital stock ($k$) or labor ($l$) for most economies.}
  \item{Figure~\ref{fig:CIvsParam_Lattice_Graph} indicates that energy (in all quantifications: $q$, $x$, and $u$) is complementary to capital stock ($k$) and labor ($l$) for producing economic growth, according to the CES model.}
  \item{Figures~\ref{fig:LINEXq_Params_Graph}--\ref{fig:LINEXu_Params_Graph} indicate that LINEX model fitting parameters $a_0$ and $c_t$ are significant (at the 95\% confidence level) for nearly all countries. Figure~\ref{fig:LINEX_GDP_Lattice_Graph} shows that the LINEX production function provides reasonably good fits to historical data. This evidence indicates that the LINEX production function (with energy ($e$) as its only factor of production) is a reasonable replacement the Cobb-Douglas and CES production functions which have capital stock ($k$) and labor ($l$) as factors of production.}
  \item{**** Discuss the AIC values for each of the model types. It could happen that the LINEX models have a higher AIC value than the CD or CES models. If so, that would mean that the LINEX model is actually preferred to the CD or CES models, primarily because it does a good-enough job of predicting economic output with fewer adjustable parameters (2) compared to CD (2), CDe (3), CES (4), CESe (6). Discuss with Randy Pruim during Summer 2013.}
\end{itemize}

@@

We claim that statistical properties (e.g., covariances) of historical economic data and statistical properties (e.g., confidence intervals) of production function fitting parameters are extremely useful for interpreting historical economic data and growth model results. We know of no previous studies on the role of energy in the economy that presented these statistical properties, and we presume that energy-augmentation of production functions has been done in the literature without regard to whether it is statistically meaningful to do so. That different authors can find good correlation between model predictions and historical data with very different growth model coefficients is a result of the heretofore unseen large confidence intervals on the fitting parameters. As a consequence, economic growth models that assume very different relationships among capital stock, labor, and energy can all be successful in matching historical data. In short, the production function is a canvas upon which investigators can paint their preconceived vision of the importance (or not) of energy for economic growth.

@@

<<Print_Covariance_Tables, results='asis', eval=TRUE>>=
#I wish I could say the following, but I get bogus "$US" strings appearing in the LaTeX output.
#lapply(countryAbbrevs, printCovarTable)
printCovarTable(countryAbbrev="US"); printCovarTable(countryAbbrev="UK"); printCovarTable(countryAbbrev="JP")
printCovarTable(countryAbbrev="CN"); printCovarTable(countryAbbrev="ZA")
printCovarTable(countryAbbrev="SA"); printCovarTable(countryAbbrev="IR")
printCovarTable(countryAbbrev="TZ"); printCovarTable(countryAbbrev="ZM")
@

>>>>

Values of $\varepsilon_i$ and $\phi_i$ were calculated from 
the REXS database supplied by \citet{Warr:2012a}. 
Values of $X$ were calculated by methods discussed in 
Section~\ref{sec:X_Data}. For most countries, values of 
$\phi_i$ are not constant with time, as economic sectors with 
different distributions of raw exergy inputs grow at different rates. 
\citeapos{Warr:2012a} REXS database contains information for 
developed economies (United States, United Kingdom, and Japan) 
from 1900--2000, and we used data from the 1980--2000 timeframe. 
Extending the REXS database to the BRICS, petrol, and LDC economies 
is beyond the scope of this work, because longitudinal data for 
the efficiency of different end uses ($\varepsilon_i$) and 
the distribution of exergy sources to end uses 
($\phi_i$) are extremely difficult to obtain.

@@

Each of the economic growth models discussed in 
Section~\ref{sec:Prod_Func_Forms} has parameters which must be estimated 
for each economy (see Section~\ref{sec:Economy_Types}) and 
each energy quantification (see Section~\ref{sec:Quantification_of_Energy}) 
via least-squares methods. The following subsections discuss 
parameter estimation methods and interpret the results.

@@

% \citet{Henningsen:2011td}, in their detailed analysis of 
% \citet{Kemfert:1998vd}, found that 
% 
% \begin{quote}
% \ldots the Levenberg-Marquardt and the PORT algorithms 
% are---at least in this study---most likely to find the 
% coefficients that give the best fit to the model, 
% where the PORT algorithm can be used to restrict the estimates 
% to the economically meaningful region.
% \end{quote}

@@

\begin{itemize}
  \item{The use of chemical exergy, as opposed to thermal energy, 
        in economic growth models is probably unnecessary. 
        Table~\ref{tab:Exergy_Allocation} shows that exergy allocation factors are 
        all close to unity. 
        Therefore, there is a small difference in absolute magnitude 
        between $Q$ and $X$. 
        However, there is no difference in indexed magnitude 
        between $q$ and $x$, unless there are large changes 
        in energy mix for an economy. 
        Figure~\ref{fig:Historical_Lattice_Graph} shows that 
        $q$ and $x$ lines lie atop one another for almost every economy.}
  \item{The concept of useful work ($u$) is potentially useful and 
        should be explored beyond developed economies. 
        Figure~\ref{fig:Historical_Lattice_Graph} shows that $u$ grows 
        at very different rates from thermal energy ($q$) or exergy ($x$). 
        Useful work may contain different information from thermal energy and exergy.}
  \item{Figure~\ref{fig:CDeGridLattice} shows that the set 
        of parameters ($\lambda$, $\alpha$, $\beta$, and $\gamma$) that 
        minimizes $SSE$ for the energy-augmented Cobb-Douglas production function 
        always (with the exception of Saudi Arabia) has 
        $\beta = 0$ or $\gamma = 0$. 
        This result implies that it is unnecessary to include both labor ($l$) 
        and energy ($e$) in the Cobb-Douglas production function.}
  \item{Partial residual plots for both the Cobb-Douglas model 
        (Figures~\ref{fig:CDPartialResidualsQ}--\ref{fig:CDPartialResidualsU} and 
        CES model (Figures~\ref{fig:CESPartialResidualsQ}--\ref{fig:CESPartialResidualsU}) 
        indicate that augmenting the Cobb-Douglas and CES production functions 
        with energy is not beneficial.}
  \item{Figure~\ref{fig:CIvsParam_Lattice_Graph} shows that important energy parameters 
        for both the energy-augmented Cobb-Douglas and CES production functions 
        ($\gamma$ and $\delta$, respectively) have large confidence intervals such that 
        we are unable to say (at the 95\% confidence level) that energy is 
        an important factor for economic growth, 
        vis-\`{a}-vis capital stock ($k$) or labor ($l$) for most economies.}
  \item{Figure~\ref{fig:CIvsParam_Lattice_Graph} indicates that 
        energy (in all quantifications: $q$, $x$, and $u$) is complementary 
        to capital stock ($k$) and labor ($l$) for producing economic growth, 
        according to the CES model.}
  \item{Figures~\ref{fig:LINEXq_Params_Graph}--\ref{fig:LINEXu_Params_Graph} 
        indicate that LINEX model fitting parameters $a_0$ and $c_t$ are 
        significant (at the 95\% confidence level) for nearly all countries. 
        Figure~\ref{fig:LINEX_GDP_Lattice_Graph} shows that 
        the LINEX production function provides reasonably good fits 
        to historical data. 
        This evidence indicates that the LINEX production function 
        (with energy ($e$) as its only factor of production) is 
        a reasonable replacement for the Cobb-Douglas and CES production functions 
        which have capital stock ($k$) and labor ($l$) as factors of production.}
  \item{**** Discuss the AIC values for each of the model types. 
        It could happen that the LINEX models have a higher AIC value than 
        the CD or CES models. 
        If so, that would mean that the LINEX model is actually preferred 
        to the CD or CES models, primarily because it does a good-enough job 
        of predicting economic output with fewer adjustable parameters (2) 
        compared to CD (2), CDe (3), CES (4), CESe (6). 
        Discuss with Randy Pruim during Summer 2013. ****}
\end{itemize}

@@

We claim that statistical properties (e.g., covariances) 
of historical economic data and statistical properties 
(e.g., confidence intervals) of production function fitting parameters are 
extremely useful for interpreting historical economic data and growth model results. 
We know of no previous studies on the role of energy in the economy that 
presented these statistical properties, 
and we presume that energy-augmentation of production functions has been done 
in the literature without regard to whether it is statistically meaningful to do so. 
That different authors can find good correlation 
between model predictions and historical data 
with very different growth model coefficients is 
a result of the heretofore unseen large confidence intervals 
on the fitting parameters. 
As a consequence, economic growth models that assume very different relationships 
among capital stock, labor, and energy can all be successful in matching historical data. 
In short, the production function becomes a canvas upon which 
investigators can paint their preconceived vision of the importance 
(or not) of energy for economic growth.

@@

<<Print_Covariance_Tables, results='asis', eval=TRUE>>=
#I wish I could say the following, but I get bogus "$US" strings appearing in the LaTeX output.
#lapply(countryAbbrevs, printCovarTable)
printCovarTable(countryAbbrev="US"); printCovarTable(countryAbbrev="UK")
printCovarTable(countryAbbrev="JP")
printCovarTable(countryAbbrev="CN"); printCovarTable(countryAbbrev="ZA")
printCovarTable(countryAbbrev="SA"); printCovarTable(countryAbbrev="IR")
printCovarTable(countryAbbrev="TZ"); printCovarTable(countryAbbrev="ZM")
@
